---
title: "hw1"
author: "Amanda Howarth"
date: "2/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(stargazer) 
library(leaps)
library(FNN)
library(ModelMetrics)
library(caret)
library(boot)
library(Rcpp)
library(microbenchmark)
library(ISLR)
library(glmnet)
library(corrplot)
library(plotmo)
```

#should everything be a factor variable (in matrix all variables have to be numeric; removes factor values and turns them into 1 and 0 (e.g. league, division))

```{r}
test =
  read_excel(path = "./data/solubility_test.xlsx", sheet = 1) %>% 
  janitor::clean_names() %>% 
  na.omit()

train =
  read_excel(path = "./data/solubility_train.xlsx", sheet = 1) %>% 
  janitor::clean_names() %>% 
  na.omit()
```

#QUSETION 1
1. (a) Fit a linear model using least squares on the training data and calculate the mean square error using the test data.

#not sure you need [-tRows]
```{r}
ctrl1 <- trainControl(method = "repeatedcv", number = 10, repeats = 5)

set.seed(2)
lmfit <- train(solubility~., 
                data = train, 
                method = "lm", 
                trControl = ctrl1)
lmfit

#test error 
pred_lm <- predict(lmfit, test)
mse(test$solubility, pred_lm)

```
test error (MSE) =0.5558898

#QUSETION 2
1b. Fit a ridge regression model on the training data, with λ chosen by cross-validation. Report the test error.
#check the sequence
```{r}
x <- model.matrix(solubility~.,train)[,-1]
y <- train$solubility

ctrl1 <- trainControl(method = "repeatedcv", number = 10, repeats = 5)

set.seed(2)
ridge.fit <- train(x, y,
                   method = "glmnet",
                   tuneGrid = expand.grid(alpha = 0, 
                                          lambda = exp(seq(-1, 10, length=100))), 
                   preProc = c("center", "scale"),
                   trControl = ctrl1)
ridge.fit

plot(ridge.fit, xTrans = function(x) log(x))

#lambda value 
ridge.fit$bestTune 

#model coefficients 
coef(ridge.fit$finalModel,ridge.fit$bestTune$lambda) 

#test error 
pred_ridge <- predict(ridge.fit, test)
mse(test$solubility, pred_ridge)
```

"Tuning parameter 'alpha' was held constant at a value of 0
RMSE was used to select the optimal model using the smallest value.
The final values used for the model were alpha = 0 and lambda = 0.3678794."

lambda = 0.3678794
test error= 0.545737


1c. Fit a lasso model on the training data, with λ chosen by cross-validation. Report the test error, along with the number of non-zero coefficient estimates.
```{r}
set.seed(2)
lasso.fit <- train(x, y,
                   method = "glmnet",
                   tuneGrid = expand.grid(alpha = 1, 
                                          lambda = exp(seq(-1, 10, length=100))),
                   preProc = c("center", "scale"),
                   trControl = ctrl1)
lasso.fit

plot(lasso.fit, xTrans = function(x) log(x))

#lamda value 
lasso.fit$bestTune

#model coefficients 
coef(lasso.fit$finalModel,lasso.fit$bestTune$lambda)

#test error 
pred_lasso <- predict(lasso.fit, test)
mse(test$solubility, pred_lasso)
```
lambda = 0.3678794
test error= 1.328538
#is getting the same lambda as the ridge.fit weird?

1d. Fit a principle component regression model on the training data, with M chosen by cross-validation. Report the test error, along with the value of M selected by cross-validation.
```{r}

ctrl1 <-trainControl(method = "repeatedcv", number = 10, repeats = 5)

set.seed(2)
pcr.fit <-train(x, y,
                method = "pcr",
                tuneGrid  =data.frame(ncomp = 1:228), 
                trControl = ctrl1,
               preProc =c("center", "scale"))
pcr.fit 

#M value
pcr.fit$bestTune

#test error 
pred_pcr <- predict(pcr.fit, test)
mse(test$solubility, pred_pcr)

ggplot(pcr.fit, highlight = TRUE) + theme_bw()

```
M = 157
MSE = 0.549917


1e. Briefly discuss the results obtained in (a)∼(d).

1f. Which model will you choose for predicting solubility?
```{r, fig.width=5}
resamp <- resamples(list(lm = lmfit,
                         ridge = ridge.fit, 
                         lasso = lasso.fit, 
                         pcr = pcr.fit))

summary(resamp)
```

try to get to pdf!!


